A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-15 text/formal-defnition.tex
Revised until Section 3.1.

\AI{Rename ``type variables'' to ``type-level constants''. }
We assume the denumerable set of type-level constants, ranged over by
metavariables \(X, Y, Z\), the denumerable set of variables, ranged
over by \(x,y,z\), the denumerable set of constants, ranged over by
\(c\), and the denumerable set of stage variables, ranged over by
\(\alpha, \beta, \gamma\).  The metavariables \(A, B, C\) range over
sequences of stage variables; we write \(\varepsilon\) for the empty
sequence.  \AI{I prefer $\varepsilon$ for the nonempty sequence.}  \LMD is defined by the following grammar:

\AI{``Transition'' is still used.}

\begin{align*}
    % \textrm{Type variables}  &   &                          & X,Y,Z                                                                                                      \\
    % \textrm{Variables}       &   &                          & x,y,z                                                                                                      \\
    % \textrm{Stage variables} &   &                          & \alpha,\beta,\gamma                                                                                        \\
    % \textrm{Stage}           &   &                          & A,B,C                                                                                                      \\
    \textrm{Kinds}           &   & K,J,I,H,G                & ::= * \mid \Pi x:\tau.K                                                                                    \\
    \textrm{Types}           &   & \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau                          \\
%     \textrm{Constants}       &   &                          & c                                                                                                          \\
    \textrm{Terms}           &   & M,N,L,O,P                & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M                                            \\
                             &   &                          & \ \ \ \ \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \varepsilon \mid \%_\alpha M                          \\  
    \textrm{Signature}       &   & \Sigma                   & ::= \emptyset \mid \Sigma, X::K \mid \Sigma, c:\tau                                                                        \\
    \textrm{Contexts}        &   & \Gamma                   & ::= \emptyset \mid  \Gamma,x:\tau @A\ (x\not\in\textrm{Domain}(\G)\ \text{and}\ x\not\in\textrm{FV}(\tau)) \\
\end{align*}

\AI{Add $M\,\alpha$ to terms.}

@@

A kind, which is used to classify types, is either $*$, the kind of
proper types (types that terms inhabit), or $\Pi x\colon\tau.K$, the kind
of type operators that takes $x$ as an argument of type $\tau$ and returns a type
of kind $K$.
% Types
% of terms have kind $*$ and dependent types have $\Pi$-kinds.  For
% example, $\lambda x:\I.x$ has type $\Pi x:\I.\I$, which has $*$ kind.
% \red{この段落は短いので型の段落と結合するか?}
% Types
A type is a type-level constant $X$, which is declared in the signature with its kind, a dependent function type $\Pi x:\tau_1.\tau_2$,
an application of a type (operator) of $\Pi$-kind to a term, a code type $\TW_\alpha$, or an $\alpha$-closed type $\F\alpha.\tau$.
We sometimes abbreviate $\Pi x:\tau_1.\tau_2$ to $\tau_1 \rightarrow \tau_2$ if
$x$ is not a free variable in $\tau_2$. \AI{This should be mentioned after free variables are introduced.}
An application of a term to a dependent function type \AI{this sounds like $M\, (\Pi x:\tau_1.\tau_2)$} is, for example, $\text{Vector}\ 10$
if type-level constant $\text{Vector}$ has the kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ denotes a code fragment of a term of type $\tau$.
An $\alpha$-closed type corresponds to a runnable code fragment.
Type constructors tie strongly \AI{Is ``tie strongly'' standard?} in the order of 
an application of a term to a dependent function type $\tau\ M$,
a code type $\TW_{\alpha} \tau$,
a dependent function type $\Pi x:\tau.\tau$, 
an $\alpha$-closed type $\F\alpha.\tau$.
Therefore, $\F\alpha.\TW_{\alpha} \Pi x:\I.\text{Vector}\ 5$ is interpreted as
$\F\alpha.(\TW_{\alpha} (\Pi x:\I.(\text{Vector}\ 5)))$. \AI{I don't understand this rule...}

@@

Terms include ordinary (explicitly typed) \(\lambda\)-terms, constants,
whose type is declared in signature $\Sigma$, and the following five forms
related to multi-stage programming:
$\TB_\alpha M$ represents a code fragment; $\TB_\alpha M$ represents escape;
$\Lambda\alpha.M$ is a stage variable abstraction;
$M\ \varepsilon$ is an application of $\varepsilon$ to a stage variable \AI{I think the usage of ``application'' is wrong again}; and 
$\%_\alpha M$ is an operator for cross-stage persistence.
The prefix operators $\TB_\alpha, \TBL_\alpha$, and $\%_\alpha$ tie stronger than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.
For example, $\F\alpha.\lambda x:\I.\TB_\alpha x\ y$ means $\F\alpha.(\lambda x:\I.(\TB_\alpha x)\ y)$.

@@

We adopt a \LLF-like system, where constants and type-level constants for base types \AI{``for base types''?} are given in the signature $\Sigma$, which 
are a sequence of declarations of the form $c:\tau$ and $X:K$.
For example, when we use integers in \LMD, $\Sigma = \I :: *, 1:\I, 2:\I, \cdots$.  \AI{It sounds like we need an infinite signature...}
\AI{Change Int to Bool.}

% Free variables
As usual, the variable $x$ is bound in $\lambda x:\tau.M$
and the transition variable $\alpha$ is bound in $\Lambda \alpha.M$.
\AI{What about other binders such as $\Pi$?}
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.

% Contexts
Contexts \AI{Why not call them ``type environments'' to avoid confusion with evaluation contexts?} are sequences of triples of a variable, its type, and its stage.
$\textrm{Domain}$ is a function from a content to a set of variables defined in the context.
\AI{I'd prefer \textit{dom}.}
The side condition of $x\not\in\textrm{Domain}(\G)$ prevent multiple appearances of the same variable and
$x\not\in\textrm{FV}(\tau)$ is needed in the proof of Substitution Lemmas.

\AI{We should note that constants and type-level constants are not staged.}
\paragraph{Remark:}  \AI{Comparison with Hanada--Igarashi.}

@@

The relation $\longrightarrow_\beta$ represents ordinary $\beta-$reduction in the \(\lambda\)-calculus; the relation 
$\longrightarrow_\blacklozenge$ represents that quotation $\TB_\alpha M$ is canceled by escape and $M$ is spliced into the code fragment surrounding the escape;
the relation $\longrightarrow_\Lambda$ means that a stage abstraction applied to the empty stage $\varepsilon$ reduces to the body of abstraction
where $\varepsilon$ is substituted for the stage variable.
Application of a non-$\varepsilon$ stage to a stage abstraction is prohibited in order to simplify \LMD. \AI{This should have been mentioned earlier in the comparison with Hanada--Igarashi.}
There is no reduction rule for CSP as with Hanada and Igarashi \cite{Hanada2014}.
The CSP operator $\%_\alpha$ disappears when $\varepsilon$ is substituted to $\alpha$.
% $\TB_\alpha$ and $\TBL_\alpha$ disappears in the same way as $\%_\alpha$.

@@

In this section, we define the type system of \LMD.
% \LMD is little complicated because it contains dependent types.
It consists of six judgment forms for typing, kinding, well-formed kinding, term equivalence, type equivalence, and kind equivalence.
We list the judgments forms in Figure~\ref{fig:LMD-six-judgements}.
\AI{equality vs. equivalence?}
\AI{A few judgment forms do not match rules.}

\begin{figure}
	\begin{center}
		\begin{align*}
			\G & \V M : \tau @ A   & \text{typing} \\
			\G & \V \tau :: K @ A  & \text{kinding} \\
			\G & \V K \iskind @ A  & \text{well-formed kinding} \\
			\G & \V M \E N @ A     & \text{term equality} \\
			\G & \V \tau \E \sigma & \text{type equality} \\
			\G & \V K \E J @ A     & \text{kind equality}
		\end{align*}
		\caption{Six judgment forms of the type system of \LMD.}
		\label{fig:LMD-six-judgements}
	\end{center}
\end{figure}

\AI{Insert a tilde (which is a space without a line break) between ``Figure'' and ref.}
\begin{definition}[Typing]
  The typing relation $ \G \V M : \tau @ A $ is the least relation closed under the rules in Figure~\ref{fig:typing-rules}.
  \AI{Precisely speaking, it cannot be defined independently of other judgments.
  How is dependency between judgments?}
\end{definition}

\AI{``judgment'' vs. ``judgement''.}  We start from defining typing of
\LMD.  Typing rules of \LMD are shown in Figure
\ref{fig:typing-rules}.  The rule \TConst{} means any constants in the
signature can appear at any stage.  \AI{This should be discussed
  earlier.}  For example, if we have a signature $\Sigma$ which is
$\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}:
\textrm{bool}$, the derivation tree in
Figure~\ref{fig:tconst-derivation-tree} is admissible.  \AI{Don't we
  need signature well formedness?  If $c:\tau \in \Sigma$, then $\tau$
  should be a well-formed type under ... what?}  The rules \TVar,
\TAbs, and \TApp{} are almost the same as those in the simply typed
lambda calculus.  \AI{... or \(\lambda\)LF?  We are working in a dependently type system...}  Additional conditions are that subterms must be
typed at the same stage (\TAbs{} and \TApp); the type
annotation/declaration on a variable has to be a proper type of kind
$*$ (\TAbs) at the stage where it is declared (\TVar{} and \TAbs).

@@

% Typing rules for a multi-stage calculus
\AI{This paragraph is not very informaticve.  We shouldn't just refer redears to previous work.}
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp{} are rules for a multi-stage calculus.
They are corresponding to brackets, escape, making a stage abstraction, 
application of $\varepsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

@@

Interestingly, Hanada and Igarashi~\cite{} rejected the idea of
reduction to remove $\%_\alpha$ when they developed \LTP{}, as such
reduction does not match the operational behavior of the CSP operator
in implementation.  However, as an equational system for multi-stage
programs, the rule \QPercent makes sense and as we argue \AI{where?}
it is practically significant.

@@

\begin{definition}[Evaluation Context and Redex]
  The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$,
the sets of $\varepsilon$-redexes (ranged over by $R^\varepsilon$) and $\alpha$-redexes (ranged over by $R^\alpha$) are defined by the following grammar.
	In the grammar, $A$ is assumed to be nonempty (but $B,A'$ can be empty).
	\begin{align*}
		E^\varepsilon_B \in ECtx^\varepsilon_B & ::= \square\ (\text{if\ } B = \varepsilon) \mid E^\varepsilon_B\ M \mid v^e\ E^\varepsilon_B     
		\mid \TB_\alpha E^\alpha_B \mid \Lambda\alpha.E^\varepsilon_B
		\mid E^\varepsilon_B\ \varepsilon                                                                                                 \\
		E^A_B \in ECtx^A_B               & ::= \square\ (\text{if } A = B) \mid \lambda x:\tau.E^A_B \mid E^A_B\ M \mid v^A\ E^A_B 
		\mid E^\varepsilon_B \mid \TB_\alpha E^{A\alpha}_B
		\mid \TBL_\alpha E^{A'}_B \ (\text{where } A'\alpha = A)                                                                    \\
		                                 & \quad \mid \Lambda\alpha.E^\varepsilon_B                                                   
		\mid E^A_B\ \varepsilon \mid \%_\alpha\ E^{A'}_B \ (\text{where } A'\alpha = A)                                                \\
\\
		 R^\varepsilon  & ::= (\lambda x:\tau.M)\ v^\varepsilon \mid (\Lambda\alpha.v^\varepsilon)\ \varepsilon                                       & \\
		 R^\alpha    & ::=\ \TBL_\alpha \TB_\alpha M                                                                                      & \\	\end{align*}
\end{definition}

@@

\begin{definition}[Staged Reduction]
	The staged reduction relation, written $M \longrightarrow_s M'$, is defined by 
	the least relation closed under the rules below.
% \begin{figure}
% 	\begin{center}
		\begin{align*}
			E^A_\varepsilon [(\lambda x:\tau.M)\ v^\varepsilon]       & \longrightarrow_s E^A_\varepsilon[M[x\mapsto v^\varepsilon]]             \\
			E^A_\varepsilon [(\Lambda\alpha.v^\varepsilon)\ \varepsilon] & \longrightarrow_s E^A_\varepsilon[v^\varepsilon[\alpha\mapsto \varepsilon]] \\
			E^A_\alpha [\TBL_\alpha \TB_\alpha v^\alpha]        & \longrightarrow_s E^A_\alpha[v^\alpha]                             \\
		\end{align*}
     %    	\caption{Staged Reduction}
     %    	\label{fig:staged-reduction}
     %    \end{center}
     % \end{figure}
\end{definition}

>>>>

% \AI{Rename ``type variables'' to ``type-level constants''. }
% \AI{I prefer $\varepsilon$ for the nonempty sequence.}  
We assume the denumerable set of type-level constants, ranged over by
metavariables \(X, Y, Z\), the denumerable set of variables, ranged
over by \(x,y,z\), the denumerable set of constants, ranged over by
\(c\), and the denumerable set of stage variables, ranged over by
\(\alpha, \beta, \gamma\).  The metavariables \(A, B, C\) range over
sequences of stage variables; we write \(\varepsilon\) for the empty
sequence. \LMD is defined by the following grammar:

% \AI{``Transition'' is still used.}

\begin{align*}
	% \textrm{Type variables}  &   &                          & X,Y,Z                                                                                                      \\
	% \textrm{Variables}       &   &                          & x,y,z                                                                                                      \\
	% \textrm{Stage variables} &   &                          & \alpha,\beta,\gamma                                                                                        \\
	% \textrm{Stage}           &   &                          & A,B,C                                                                                                      \\
	\textrm{Kinds}             &  & K,J,I,H,G                & ::= * \mid \Pi x:\tau.K                                                           \\
	\textrm{Types}             &  & \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
	%     \textrm{Constants}       &   &                          & c                                                                                                          \\
	\textrm{Terms}             &  & M,N,L,O,P                & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M                   \\
	                           &  &                          & \ \ \ \ \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ A \mid \%_\alpha M        \\
	\textrm{Signature}         &  & \Sigma                   & ::= \emptyset \mid \Sigma, X::K \mid \Sigma, c:\tau                               \\
	\textrm{Type environments} &  & \Gamma                   & ::= \emptyset \mid  \Gamma,x:\tau @A                                              \\
\end{align*}

% \AI{Add $M\,\alpha$ to terms.}

@@

A kind, which is used to classify types, is either $*$, the kind of
proper types (types that terms inhabit), or $\Pi x\colon\tau.K$, the kind
of type operators that takes $x$ as an argument of type $\tau$ and returns a type
of kind $K$.
% Types
% of terms have kind $*$ and dependent types have $\Pi$-kinds.  For
% example, $\lambda x:\I.x$ has type $\Pi x:\I.\I$, which has $*$ kind.
% \red{この段落は短いので型の段落と結合するか?}
% Types
A type is a type-level constant $X$, which is declared in the signature with its kind, a dependent function type $\Pi x:\tau_1.\tau_2$,
an application of a term to a type (operator) of $\Pi$-kind, a code type $\TW_\alpha$, or an $\alpha$-closed type $\F\alpha.\tau$.
An application of a term to a type (operator) of $\Pi$-kind is, for example, $\text{Vector}\ 10$
if type-level constant $\text{Vector}$ has the kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ denotes a code fragment of a term of type $\tau$.
An $\alpha$-closed type corresponds to a runnable code fragment.

% Type constructors tie strongly \AI{Is ``tie strongly'' standard?} in the order of 
An application of a term to a type (operator) of $\Pi$-kind: $\tau\ M$ connect tighter than
the type constructor $\TW_{\alpha}$ and $\TW_{\alpha}$ connect tighter than
$\Pi$ in dependent types such as $\Pi x:\tau.\tau$ and $\Pi$ connect tighter than
$\F$ in types for stage abstraction such as $\F\alpha.\tau$.
Therefore, $\F\alpha.\TW_{\alpha} \Pi x:\I.\text{Vector}\ 5$ is interpreted as
$\F\alpha.(\TW_{\alpha} (\Pi x:\I.(\text{Vector}\ 5)))$.
% \AI{I don't understand this rule...}

@@

Terms include ordinary (explicitly typed) \(\lambda\)-terms, constants,
whose type is declared in signature $\Sigma$, and the following five forms
related to multi-stage programming:
$\TB_\alpha M$ represents a code fragment; $\TB_\alpha M$ represents escape;
$\Lambda\alpha.M$ is a stage variable abstraction;
$M\ A$ is an application of stage $A$ to a stage abstraction; and
$\%_\alpha M$ is an operator for cross-stage persistence.
The prefix operators $\TB_\alpha, \TBL_\alpha$, and $\%_\alpha$ connect tighter than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.
For example, $\F\alpha.\lambda x:\I.\TB_\alpha x\ y$ means $\F\alpha.(\lambda x:\I.(\TB_\alpha x)\ y)$.
% \AI{I think the usage of ``application'' is wrong again}

@@

We adopt a \LLF-like system, where constants and type-level constants are given in the signature $\Sigma$, which
are a sequence of declarations of the form $c:\tau$ and $X:K$.
For example, when we use integers in \LMD, $\Sigma = \B :: *, true:\B, false:\B$.
Please note that constants or type-level constants has no stage.
This means that we can use them at any stage.
% Type environments
Type environments are sequences of triples of a variable, its type, and its stage.
We define well-formed signature and well-formed type environments later
in order to prove theorems.

% Free variables
As usual, the variable $x$ is bound in $\lambda x:\tau.M$ or $\Pi x:\tau.\tau$ and
and the stage variable $\alpha$ is bound in $\Lambda \alpha.M$ or $\F\alpha.\tau$.
We sometimes abbreviate $\Pi x:\tau_1.\tau_2$ to $\tau_1 \rightarrow \tau_2$ if
$x$ is not a free variable in $\tau_2$.
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.
% \AI{This should be mentioned after free variables are introduced.}
% \AI{What about other binders such as $\Pi$?}

% \AI{We should note that constants and type-level constants are not staged.}
\paragraph{Remark:}  \AI{Comparison with Hanada--Igarashi.}

@@

The relation $\longrightarrow_\beta$ represents ordinary $\beta-$reduction in the \(\lambda\)-calculus; the relation
$\longrightarrow_\blacklozenge$ represents that quotation $\TB_\alpha M$ is canceled by escape and $M$ is spliced into the code fragment surrounding the escape;
the relation $\longrightarrow_\Lambda$ means that a stage abstraction applied to the empty stage $\varepsilon$ reduces to the body of abstraction
where $\varepsilon$ is substituted for the stage variable.
There is no reduction rule for CSP as with Hanada and Igarashi \cite{Hanada2014}.
The CSP operator $\%_\alpha$ disappears when $\varepsilon$ is substituted to $\alpha$.
% $\TB_\alpha$ and $\TBL_\alpha$ disappears in the same way as $\%_\alpha$.

@@

In this section, we define the type system of \LMD.
% \LMD is little complicated because it contains dependent types.
It consists of six judgment forms for typing, kinding, well-formed kinding, term equivalence, type equivalence, and kind equivalence.
We list the judgments forms in Figure~\ref{fig:LMD-six-judgments}.
\AI{equality vs. equivalence?}
\AI{A few judgment forms do not match rules.}

\begin{figure}
	\begin{center}
		\begin{align*}
			\G & \V M : \tau @ A       & \text{typing}              \\
			\G & \V \tau :: K @ A      & \text{kinding}             \\
			\G & \V K \iskind @ A      & \text{well-formed kinding} \\
			\G & \V M \E N @ A         & \text{term equivalence}    \\
			\G & \V \tau \E \sigma @ A & \text{type equivalence}    \\
			\G & \V K \E J @ A         & \text{kind equivalence}
		\end{align*}
		\caption{Six judgment forms of the type system of \LMD.}
		\label{fig:LMD-six-judgments}
	\end{center}
\end{figure}

\AI{Insert a tilde (which is a space without a line break) between ``Figure'' and ref.}
\begin{definition}[Typing, Kinding, Well-formed Kinding, Term Equivalence, Type Equivalence, Kind Equivalence]
	The typing relation $ \G \V M : \tau @ A $,
	the kinding relation $\G \V \tau :: K @ A$,
	the well-formed kinding relation $\G \V K \iskind @ A$,
	the term equivalence relation $\G \V M \E N @ A$,
	the type equivalence relation $\G \V \tau \E \sigma @ A$,
	the kind equivalence relation $\G \V K \E J @ A$,
	are the least relation closed under the rules in Figure~\ref{fig:typing-rules}.
\end{definition}
This definition is little complicated to define six relations at the same time.
We canot define them one by one because they are dependent on each other.

We introduce a well-formed signature and well-formed type environments.
$\textit{dom}$ is a function from a content to a set of variables defined in the context.
\begin{definition}[Well-formed Signature]
	\begin{itemize}
		\item $\emptyset$ is a well-formed signature.
		\item if $\Sigma$ is a well-formed signature and $\V K :: * @ \varepsilon$ and $X\notin\textit{dom}(\Sigma)$ \\
		      then $\Sigma,X::K$ is a well-formed signature.
		\item if $\Sigma$ is a well-formed signature and $\V \tau :: K @ \varepsilon$ and $c\notin\textit{dom}(\Sigma)$ \\
		      then $\Sigma,c:\tau$ is a well-formed signature.
	\end{itemize}
\end{definition}
\begin{definition}[Well-formed Type Environments]
	\begin{itemize}
		\item $\emptyset$ are well-formed type environments.
		\item if $\G$ are well-formed type environments and $x\not\in\textit{dom}(\G)$ and $x\not\in\textrm{FV}(\tau)$ \\
			  then $\G,x:\tau@A$ are well-formed type environments.
	\end{itemize}
\end{definition}
From here, we assume that all signature and type environments are well-formed.

 We start from defining typing of
\LMD.  Typing rules of \LMD are shown in Figure
\ref{fig:typing-rules}.  The rule \TConst{} means any constants in the
signature can appear at any stage.  \AI{This should be discussed
	earlier.}  For example, if we have a signature $\Sigma$ which is
$\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}:
	\textrm{bool}$, the derivation tree in
Figure~\ref{fig:tconst-derivation-tree} is admissible.  \AI{Don't we
	need signature well formedness?  If $c:\tau \in \Sigma$, then $\tau$
	should be a well-formed type under ... what?}  The rules \TVar,
\TAbs, and \TApp{} are almost the same as those in the simply typed
lambda calculus.  \AI{... or \(\lambda\)LF?  We are working in a dependently type system...}  Additional conditions are that subterms must be
typed at the same stage (\TAbs{} and \TApp); the type
annotation/declaration on a variable has to be a proper type of kind
$*$ (\TAbs) at the stage where it is declared (\TVar{} and \TAbs).

@@

% Typing rules for a multi-stage calculus
\AI{This paragraph is not very informaticve.  We shouldn't just refer redears to previous work.}
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp{} are rules for a multi-stage calculus.
They are corresponding to brackets, escape, making a stage abstraction,
application of stage $A$ to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

@@

Interestingly, Hanada and Igarashi~\cite{Hanada2014} rejected the idea of
reduction to remove $\%_\alpha$ when they developed \LTP{}, as such
reduction does not match the operational behavior of the CSP operator
in implementation.  However, as an equational system for multi-stage
programs, the rule \QPercent makes sense and as we argue next
it is practically significant.

@@

\begin{definition}[Evaluation Context and Redex]
	The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$,
	the sets of $\varepsilon$-redexes (ranged over by $R^\varepsilon$) and $\alpha$-redexes (ranged over by $R^\alpha$) are defined by the following grammar.
	In the grammar, $A$ is assumed to be nonempty (but $B,A'$ can be empty).
	\begin{align*}
		E^\varepsilon_B \in ECtx^\varepsilon_B & ::= \square\ (\text{if\ } B = \varepsilon) \mid E^\varepsilon_B\ M \mid v^e\ E^\varepsilon_B
		\mid \TB_\alpha E^\alpha_B \mid \Lambda\alpha.E^\varepsilon_B
		\mid E^\varepsilon_B\ \varepsilon                                                                                                       \\
		E^A_B \in ECtx^A_B                     & ::= \square\ (\text{if } A = B) \mid \lambda x:\tau.E^A_B \mid E^A_B\ M \mid v^A\ E^A_B
		\mid E^\varepsilon_B \mid \TB_\alpha E^{A\alpha}_B
		\mid \TBL_\alpha E^{A'}_B \ (\text{where } A'\alpha = A)                                                                                \\
		                                       & \quad \mid \Lambda\alpha.E^\varepsilon_B
		\mid E^A_B\ \varepsilon \mid \%_\alpha\ E^{A'}_B \ (\text{where } A'\alpha = A)                                                         \\
		\\
		R^\varepsilon                          & ::= (\lambda x:\tau.M)\ v^\varepsilon \mid (\Lambda\alpha.v^\varepsilon)\ \varepsilon        & \\
		R^\alpha                               & ::=\ \TBL_\alpha \TB_\alpha M                                                                & \\\end{align*}
\end{definition}

@@

\begin{definition}[Staged Reduction]
	The staged reduction relation, written $M \longrightarrow_s M'$, is defined by
	the least relation closed under the rules below.
	% \begin{figure}
	% 	\begin{center}
	\begin{align*}
		E^A_\varepsilon [(\lambda x:\tau.M)\ v^\varepsilon] & \longrightarrow_s E^A_\varepsilon[M[x\mapsto v^\varepsilon]]      \\
		E^A_\varepsilon [(\Lambda\alpha.v^\varepsilon)\ A]  & \longrightarrow_s E^A_\varepsilon[v^\varepsilon[\alpha\mapsto A]] \\
		E^A_\alpha [\TBL_\alpha \TB_\alpha v^\alpha]        & \longrightarrow_s E^A_\alpha[v^\alpha]                            \\
	\end{align*}
	%    	\caption{Staged Reduction}
	%    	\label{fig:staged-reduction}
	%    \end{center}
	% \end{figure}
\end{definition}
